---
title: "Race voting blocs sensitivity tests"
author:
  - ""
date: ""
output:
  html_document:
    df_print: paged
    toc_depth: 2
    theme: paper
    highlight: tango
editor_options:
  chunk_output_type: inline
---

```{r setup, include = FALSE}
knitr::opts_knit$set(root.dir = here::here())
```

# Setup

```{r packages, warning=FALSE, message=FALSE}
library(dplyr)
library(tidyr)
library(readr)
library(stringr)
library(ggplot2)
library(cowplot)
library(kableExtra)
```

Read in results from race voting bloc analyses. Also read in McDonald's turnout-by-race estimates.

```{r}
vbdf_cps  <- readRDS("data/vbdf_race_cps.rds")
vbdf_cces <- readRDS("data/vbdf_race_cces.rds")
vbdf_cces_reported   <- readRDS("data/vbdf_race_cces_reported.rds")
vbdf_anes <- readRDS("data/vbdf_race_anes.rds")
# acs-cces-cces
vbdf_acc <- readRDS("data/vbdf_race_acc.rds")
# acs-anes-anes
vbdf_aaa <- readRDS("data/vbdf_race_aaa.rds")
# main results: cps-anes-anes
vbdf_caa <- readRDS("data/vbdf_race_caa.rds")

mcdonald_turn <- read_csv("data/McDonaldTurnoutNumbers.csv")
```



## Clean up results

Recode "hispanic" (term from surveys) to "latino" (term for paper).
```{r}
vbdf_cps <- 
    mutate(vbdf_cps, 
           race = recode_factor(race, hispanic = "latino"))
vbdf_cces <-
    mutate(vbdf_cces, 
           race = recode_factor(race, hispanic = "latino"))
vbdf_cces_reported <- 
    mutate(vbdf_cces_reported, 
           race = recode_factor(race, hispanic = "latino"))
vbdf_anes <-
    mutate(vbdf_anes, 
           race = recode_factor(race, hispanic = "latino"))
vbdf_acc <- 
    mutate(vbdf_acc, 
           race = recode_factor(race, hispanic = "latino"))
vbdf_aaa <- 
    mutate(vbdf_aaa, 
           race = recode_factor(race, hispanic = "latino"))
vbdf_caa <- 
    mutate(vbdf_caa, 
           race = recode_factor(race, hispanic = "latino"))
```


Restrict to recent years and code `NA` values as other race.

```{r results = "hide"}
vbdf_caa <-
    filter(vbdf_caa, year > 1975) %>%
    mutate(year = factor(year, ordered = TRUE),
           race = if_else(is.na(race), "other", as.character(race)),
           group = race %>% tools::toTitleCase() %>% factor())


vbdf_cps <-
    filter(vbdf_cps, year > 1975) %>%
    mutate(year = factor(year, levels = levels(vbdf_caa$year), ordered = TRUE),
           race = if_else(is.na(race), "other", as.character(race)),
           group = race %>% tools::toTitleCase() %>% factor())


vbdf_cces <-
    filter(vbdf_cces, year > 1975) %>%
    mutate(year = factor(year, levels = levels(vbdf_caa$year), ordered = TRUE),
           race = if_else(is.na(race), "other", as.character(race)),
           group = race %>% tools::toTitleCase() %>% factor())

vbdf_cces_reported <-
    filter(vbdf_cces_reported, year > 1975) %>%
    mutate(year = factor(year, levels = levels(vbdf_caa$year), ordered = TRUE),
           race = if_else(is.na(race), "other", as.character(race)),
           group = race %>% tools::toTitleCase() %>% factor())


vbdf_anes <-
    filter(vbdf_anes, year > 1975) %>%
    mutate(year = factor(year, levels = levels(vbdf_caa$year), ordered = TRUE),
           race = if_else(is.na(race), "other", as.character(race)),
           group = race %>% tools::toTitleCase() %>% factor())


vbdf_aaa <-
    filter(vbdf_aaa, year > 1975) %>%
    mutate(year = factor(year, levels = levels(vbdf_caa$year), ordered = TRUE),
           race = if_else(is.na(race), "other", as.character(race)),
           group = race %>% tools::toTitleCase() %>% factor())

vbdf_acc <-
    filter(vbdf_acc, year > 1975) %>%
    mutate(year = factor(year, levels = levels(vbdf_caa$year), ordered = TRUE),
           race = if_else(is.na(race), "other", as.character(race)),
           group = race %>% tools::toTitleCase() %>% factor())
```


# CPS-ANES-ANES Results {.tabset}

```{r}
theme_set(
    theme_light() + theme(axis.text = element_text(size = 12), axis.title = element_text(size = 14))
)
```

## Composition
```{r}
p_comp <-
    vbdf_caa %>%
    mutate(year = as.numeric(as.character(year))) %>% 
    ggplot() +
    aes(x = year, y = prob_mean, shape = group, lty = group, group = group) +
    geom_line() +
    geom_point() +
    scale_shape_manual(values = c(19, 17, 15, 3)) +
    labs(title = "Composition", color = "Race", x = "Year", y = "Proportion",
         group = NULL, shape = NULL, lty = NULL) +
    scale_x_continuous(breaks = seq(1980, 2020, 10), 
                       labels = seq(1980, 2020, 10))

p_comp
```

## Turnout

```{r}
p_turnout <-
    vbdf_caa %>%
    mutate(year = as.numeric(as.character(year))) %>% 
    ggplot() +
    aes(x = year, y = pr_turnout_mean, shape = group, lty = group, group = group) +
    geom_line() +
    geom_point() +
    scale_shape_manual(values = c(19, 17, 15, 3)) +
    scale_x_continuous(breaks = seq(1980, 2020, 10), 
                       labels = seq(1980, 2020, 10)) +
    labs(title = "Turnout", color = "Race",
         x = "Year", y = "Proportion",
         group = NULL, shape = NULL, lty = NULL)

p_turnout
```

## Vote Choice

```{r}
p_votechoice <-
    vbdf_caa %>%
    mutate(year = as.numeric(as.character(year))) %>% 
    ggplot() +
    aes(x = year, group = group) +
    
    geom_ribbon(
        aes(y = pr_voterep_mean, ymin = pr_voterep_low, ymax = pr_voterep_high,
            fill = "Rep"),
        fill = "lightgrey") +
    
    geom_ribbon(
        aes(y = pr_votedem_mean, 
            ymin = pr_votedem_low, ymax = pr_votedem_high,
            fill = "Dem"),
        fill = "grey"
    ) +
    
    geom_line(aes(y = pr_voterep_mean, color = "Rep")) +
    geom_line(aes(y = pr_votedem_mean, color = "Dem")) +
    geom_hline(yintercept = 0.5, color = "grey50") +
    
    scale_color_manual(values = c(Dem = "black", Rep = "darkgrey")) +
    scale_fill_manual(values = c(Dem = "black", Rep = "darkgrey")) +
    
    facet_grid(group ~ .) +
    theme_bw() +
    # theme(legend.position = "none") +
    scale_x_continuous(breaks = seq(1980, 2020, 10), 
                       labels = seq(1980, 2020, 10)) +
    labs(title = "Vote Choice",
         x = "Year",
         y = "Pr(vote | turnout)",
         fill = NULL, color = NULL
    )

p_votechoice
```

## Net Republican Votes
```{r}
p_netrep <-
    vbdf_caa %>%
    mutate(year = as.numeric(as.character(year))) %>% 
    ggplot() +
    aes(x = year, y = net_rep_mean,
        ymin = net_rep_low, ymax = net_rep_high, group = group) +
    geom_ribbon(fill = "grey") +
    geom_point(cex = 2.5) +
    geom_hline(yintercept = 0, color = "grey50") +
    geom_line(linewidth = 1.5) +
    facet_wrap(. ~ group) +
    scale_x_continuous(breaks = seq(1980, 2020, 10), 
                       labels = seq(1980, 2020, 10)) +
    labs(y = "Net Republican Votes", x = "Year",
         group = NULL) + 
    theme_bw()

p_netrep
```


## {-}

### Write to disk
```{r}
p_grid <- 
    plot_grid(p_comp + 
                  theme(legend.position = "none", 
                        axis.text = element_text(size = 8), 
                        axis.title = element_text(size = 12)),
              p_turnout + 
                  theme(legend.position = "none", 
                        axis.text = element_text(size = 8),
                        axis.title = element_text(size = 12)),
              p_votechoice +
                  theme(legend.position = "none", 
                        axis.text = element_text(size = 8), 
                        axis.title = element_text(size = 12)), 
              rel_widths = c(1, 1.1, 1.2),
              # rel_ = c(1, 1, 1.2),
              nrow = 1)

legend_race <- 
    get_plot_component(
        p_comp + 
            guides(group = guide_legend(nrow = 1)) +
            theme(legend.position = "bottom",
                  legend.box.margin = margin(l = 70, b = 0, 0, 0)),
        "guide-box-bottom", return_all = TRUE
    )

legend_vote <- 
    get_plot_component(
        p_votechoice + 
            guides(group = guide_legend(nrow = 1)) +
            theme(legend.position = "bottom",
                  legend.box.margin = margin(l = 70, r = 0, 0, 0)),
        "guide-box-bottom", return_all = TRUE
    )

legend_grid <- 
    plot_grid(
        legend_race, legend_vote
    )

p_grid_legend <- 
    plot_grid(p_grid,
              legend_grid,
              ncol = 1, axis = "l",
              rel_heights = c(1, 0.1),
              rel_widths = c(1, 0.1)
    )

p_grid_legend
```


```{r}
ggsave(p_grid_legend, file = "docs/figures/fig1.eps", 
       height = 6, width = 10)

ggsave(p_netrep, file = "docs/figures/fig2.eps",
       height = 6, width = 10)

ggsave(p_grid_legend, file = "docs/figures/fig1.pdf", 
       height = 6, width = 10)

ggsave(p_netrep, file = "docs/figures/fig2.pdf",
       height = 6, width = 10)


ggsave(p_grid_legend, file = "docs/figures/Race_CAA_Comp.pdf", 
       height = 6, width = 10)

ggsave(p_netrep, file = "docs/figures/Race_CAA_NetRep.pdf",
       height = 6, width = 10)
```



# Sensitivity Analysis
```{r}
vb_sensitivity <-
    vb_swap <-
    function(vbdf, sens_data, bloc_var = blocs:::get_bloc_var(vbdf)){
        
        #' Sensitivity analysis by swapping estimates
        #'
        #' Test sensitivity of Net Republican Votes results to alternative estimates of density, turnout, and vote choice.
        #' Supply the data frame of voting bloc results and a similarly structured data frame of alternative bloc-level estimates.
        #' The `vb_sensitivity()` function will compute a new estimate of Net Republican Votes using any estimates contained in sens_data.
        #'
        #' @param vbdf data.frame containing baseline results of voting bloc analysis
        #' @param sens_data data.frame of alternative estimates for sensitivity analysis, stored in `vbdf` format
        
        # Checks
        require(dplyr)
        require(tidyr)
        
        blocs:::check_vbdf(vbdf)
        
        if( ! setequal(vbdf$year, sens_data$year))
            warning("Years not equal between data sets. Expect NA values.\n")
        if( ! setequal(vbdf$race, sens_data$race))
            warning("Blocs not equal between data sets. Expect NA values.\n")
        
        vbdf <- dplyr::select(vbdf, -contains("net_rep"))
        sens_data <- dplyr::select(sens_data, -contains("net_rep"))
        
        sens <-
            dplyr::left_join(vbdf, sens_data,
                             by = c("year", bloc_var),
                             # Old columns suffixed with _baseline
                             # New columns from sens_data take their place
                             # If an estimate is not present in sens_data,
                             # it isn't replaced for net_rep calculation.
                             suffix = c("_baseline", "")) %>%
            
            dplyr::mutate(net_rep_sens = prob_mean * pr_turnout_mean * (pr_voterep_mean - pr_votedem_mean) ) %>%
            dplyr::select(year, race, net_rep_sens)
        
        return(sens)
        
    }
```

Check analyses across surveys, storing results in a list.
```{r}
sens_list <-
    list(`CPS_ANES_ANES` = vbdf_caa %>%
             select(year, race, net_rep_sens = net_rep_mean))

sens_list$`ACS_CCES*_CCES` <-
    vbdf_acc %>% select(year, race, net_rep_sens = net_rep_mean)

sens_list$ACS_ANES_ANES <-
    vbdf_aaa %>% select(year, race, net_rep_sens = net_rep_mean)

sens_list$CPS_CPS_ANES <-
    vbdf_cps %>% select(year, race, net_rep_sens = net_rep_mean)

sens_list$ANES_ANES_ANES <-
    vbdf_anes %>% select(year, race, net_rep_sens = net_rep_mean)

sens_list$`CCES_CCES*_CCES` <-
    vbdf_cces %>% select(year, race, net_rep_sens = net_rep_mean)

sens_list$`CCES_CCES_CCES` <-
    vbdf_cces_reported %>% select(year, race, net_rep_sens = net_rep_mean)

```


## McDonald
Construct a data frame of alternative turnout estimates for the same years.
```{r}
mcdonald_vbdf <-
    # remove midterms
    filter(mcdonald_turn, `Turnout Rate` %in% as.character(vbdf_caa$year)) %>% 
    # construct vbdf-like object
    data.frame(
        year = factor(.$`Turnout Rate`, levels = levels(vbdf_caa$year), ordered = TRUE),
        variable = "turnout",
        white = .$`Non-Hispanic White`,
        black = .$`Non-Hispanic Black`,
        latino = .$Hispanic,
        other = .$Other
    ) %>%
    pivot_longer(
        cols = c("white", "black", "latino", "other"),
        names_to = "race",
        values_to = "pr_turnout_mean")

sens_list$ACS_McDonald_ANES <-
    vb_sensitivity(vbdf_aaa, sens_data = mcdonald_vbdf)
```

## ACS-CPS-ANES
```{r}
sens_list$ACS_CPS_ANES <-
    vb_sensitivity(vbdf_cps,
                   sens_data = vbdf_aaa %>%
                       select(year, race, contains("prob"), contains("vote")))
```

## ACS-CPS-CCES
```{r}
sens_list$ACS_CPS_ANES <-
    vb_sensitivity(vbdf_cps,
                   sens_data = vbdf_acc %>%
                       select(year, race, contains("prob"), contains("vote")))
```


## ACS-McDonald-ANES
```{r}
sens_list$ACS_McDonald_ANES <-
    vb_sensitivity(vbdf_aaa, sens_data = mcdonald_vbdf)
```

## ACS-McDonald-CCES
```{r}
sens_list$ACS_McDonald_CCES <-
    vb_sensitivity(vbdf_acc, sens_data = mcdonald_vbdf)
```

## ACS-ANES-CCES
```{r}
sens_list$ACS_ANES_CCES <-
    vb_sensitivity(vbdf_acc,
                   sens_data = select(vbdf_anes, year, race, contains("turnout")))
```

## ACS-CCES*-ANES
```{r}
sens_list$`ACS_CCES*_ANES` <-
    vb_sensitivity(vbdf_aaa,
                   sens_data = select(vbdf_cces, year, race, contains("turnout")))
```

## ACS-CCES-ANES
```{r}
sens_list$ACS_CCES_ANES <-
    vb_sensitivity(vbdf_aaa,
                   sens_data = select(vbdf_cces_reported, year, race, contains("turnout")))
```

## ACS-CPS-CCES
```{r}
sens_list$ACS_CPS_CCES <-
    vb_sensitivity(vbdf_acc,
                   sens_data = select(vbdf_cps, year, race, contains("turnout")))
```

## ACS-CPS-ANES
```{r}
sens_list$ACS_CPS_ANES <-
    vb_sensitivity(vbdf_aaa,
                   sens_data = select(vbdf_cps, year, race, contains("turnout")))
```

## CPS-ANES-CCES
```{r}
sens_list$CPS_ANES_CCES <-
    vb_sensitivity(vbdf_caa,
                   sens_data = select(vbdf_cces, year, race, contains("vote"))
    )
```



## CPS-McDonald-ANES
```{r}
sens_list$CPS_McDonald_ANES <-
    vb_sensitivity(vbdf_caa, sens_data = mcdonald_vbdf)
```

## CPS-McDonald-CCES
```{r}
sens_list$CPS_McDonald_CCES <-
    vb_sensitivity(vbdf_caa,
                   sens_data =
                       left_join(
                           mcdonald_vbdf,
                           select(vbdf_cces, year, race, contains("vote"))
                       )
    )
```

## CPS-CCES*-CCES
```{r}
sens_list$`CPS_CCES*_CCES` <-
    vb_sensitivity(vbdf_caa,
                   sens_data = select(vbdf_cces, year, race,
                                      contains("turnout"), contains("vote"))
    )
```

## CPS-CCES-CCES
```{r}
sens_list$CPS_CCES_CCES <-
    vb_sensitivity(vbdf_cps,
                   sens_data = select(vbdf_cces_reported, year, race,
                                      contains("turnout"), contains("vote"))
    )
```

## {-}

### Combine
```{r}
sens_order <-
    c("CPS_ANES_ANES", 
      "CPS_McDonald_ANES",
      "ANES_ANES_ANES", 
      "ACS_CPS_ANES", "ACS_McDonald_ANES",
      "CPS_McDonald_CCES",
      "CCES_CCES_CCES",
      "CPS_CPS_ANES", 
      "CCES_CCES*_CCES",
      "ACS_CCES*_CCES", "ACS_CCES*_ANES", "ACS_CCES_ANES", "ACS_CCES_CCES",
      "ACS_ANES_ANES", "ACS_ANES_CCES",
      "ACS_McDonald_CCES",
      "ACS_CPS_CCES",
      "CPS_CCES*_CCES", "CPS_CCES_CCES", "CPS_ANES_CCES"
    )

stopifnot(anyDuplicated(sens_order) == 0)
stopifnot(anyDuplicated(sens_list) == 0)
stopifnot(length(setdiff(names(sens_list), sens_order)) == 0)


sens_list <- sens_list[sens_order]

sens_results <-
    bind_rows(sens_list, .id = "sens_tag") %>%
    mutate(Comp = str_extract(sens_tag, "^[[:alpha:]]+(?=_)"),
           Turnout = str_extract(sens_tag, "(?<=_)[[:alpha:]*]+(?=_)"),
           Vote = str_extract(sens_tag, "(?<=_)[[:alpha:]]+$"))

```

Cast the years to columns of Net Republican Votes for each data combination and voting bloc. Split into list to print separate table for each race.
```{r}
sens_wide <-
    sens_results %>%
    pivot_wider(id_cols = c("Comp", "Turnout", "Vote", "race"),
                names_from = year,
                values_from = net_rep_sens)

sens_wide <-
    dplyr::select(sens_wide,
                  Comp, Turnout, Vote,
                  all_of(as.character(seq(1976, 2020, 4))), everything())

sens_table_list <-
    split(sens_wide, sens_wide$race) %>% lapply( \(x) select(x, -race))
```

```{r results = "asis"}
options(knitr.kable.NA = '')

sens_table_list$black %>%
    knitr::kable(digits = 3) %>%
    kableExtra::kable_styling()

sens_table_list$latino %>%
    knitr::kable(digits = 3) %>%
    kableExtra::kable_styling()

sens_table_list$white %>%
    knitr::kable(digits = 3) %>%
    kableExtra::kable_styling()

```

Write to tex file.
```{r results = "hide"}
sens_table_list$black %>%
    knitr::kable(digits = 3,
                 label = "t:black_sens",
                 caption = "Net Republican Votes from Black Citizens",
                 format = "latex", booktabs = TRUE) %>%
    kableExtra::landscape() %>%
    kableExtra::kable_styling(latex_options = "scale_down") %>%
    write("docs/tables/table_black.tex")

sens_table_list$latino %>%
    knitr::kable(digits = 3,
                 label = "t:latin_sens",
                 caption = "Net Republican Votes from Latino Citizens",
                 format = "latex", booktabs = TRUE) %>%
    kableExtra::landscape() %>%
    kableExtra::kable_styling(latex_options = "scale_down") %>%
    write("docs/tables/table_latino.tex")

sens_table_list$white %>%
    knitr::kable(digits = 3,
                 label = "t:white_sens",
                 caption = "Net Republican Votes from White Citizens",
                 format = "latex", booktabs = TRUE) %>%
    kableExtra::landscape() %>%
    kableExtra::kable_styling(latex_options = "scale_down") %>%
    write("docs/tables/table_white.tex")

```
