---
title: "Prepare CCES 2006--2018"
author:
  - "Cole Tanigawa-Lau"
date: ""
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: 3
    toc_float: yes
    theme: paper
editor_options:
  chunk_output_type: inline
---

```{r setup, include = FALSE}
knitr::opts_knit$set(root.dir = here::here())
```

```{r about, echo=FALSE, results = "asis"}
cat(
  sprintf("Updated: %s.",
          format(Sys.time(), "%b %d, %Y at %H:%M:%S", usetz = TRUE)),
  sprintf("Working directory: `%s`.", getwd()),
  sep = "\n\n"
)
```


Reads:

  - data/cces/cumulative_2006-2020.Rds

Writes:

  - data/cces_clean.rds

# Setup

```{r packages, warning=FALSE, message=FALSE}
rm(list = ls())

library(haven)
library(stringr)
library(dplyr)
library(data.table)
library(tidylog)
library(labelled)
library(questionr)
library(assertthat)
library(assertr)

source("code/functions.R")
```


## Load data

```{r}
cces_full <- readRDS("data/cces/cumulative_2006-2020.Rds")
```

Select columns: year, state, all covariates, turnout, vote choice,
```{r}
cces <-
  cces_full %>%
  filter(year %in% seq(2008, 2020, 4)) %>%
  dplyr::select(year, state = st, weight = weight_post, weight_pre = weight, case_id, citizen,
                race, gender, fincome_cces = faminc, educ, age,
                pid3, pid7, ideo5,
                voted_pres_party
                ) %>%
  as.data.table()
```


# Recode columns

```{r}
# Print label codes
select_if(cces, haven::is.labelled) %>%
  lapply(function(x) attr(x, "label"))
```


```{r}
cces <-
  mutate(cces,
         source = "CCES",
         case_id = as.character(case_id),
         year = year,
	 weight = ifelse(is.na(weight), weight_pre, weight),
         citizen = as.numeric(citizen == 1),
         south = as.numeric(state %in%
                              c("TN", "VA", "NC", "SC", "FL",
                                "GA", "AL", "MS", "LA", "AR", "TX",
                                # Schickler adds these
                                "OK", "KY")),
         race  = recode(race,
                        `1` = "white", `2` = "black", `3` = "hispanic",
                        .default = "other"),
         gender = recode(gender,
                         `1` = "male", `2` = "female",
                         .default = NA_character_),
         educ  = recode(educ,
                        `1` = "HS or less", `2` = "HS or less",
                        `3` = "some college", `4` = "some college",
                        `5` = "college", `6` = "college",
                        .default = NA_character_),
         fincome_cces = as.factor(fincome_cces),
         fincome_ordinal = as.numeric(fincome_cces),
         # RECODED FROM PID7, SEE BELOW
         # pid3 = recode(pid3,
         #               `1` = "dem",
         #               `2` = "rep",
         #               `3` = "indep",
         #               .default = "other/not sure"
         # ),
         pid7 = recode(pid7,
                       `1` = "strong dem",
                       `2` = "weak dem",
                       `3` = "indep dem",
                       `4` = "independent",
                       `5` = "indep rep",
                       `6` = "weak rep",
                       `7` = "strong rep",
                       .default = "other/not sure"),
         ideo5 = recode(ideo5,
                        `1` = "very liberal", `2` = "liberal", `3` = "moderate",
                        `4` = "conservative", `5` = "very conservative",
                        .default = "other/not sure"),
         vote_pres = recode(voted_pres_party,
                            Democratic = "dem", Republican = "rep",
                            `Other Candidate` = "other", .default = NA_character_)
  ) %>% 
  filter(citizen == 1, age >= 17)
```


Copied from Will's CCES code: age groups in 10-year bands, except under 20 is counted w/ 20s
```{r}
mfloor <- function(x,base){
  base * floor(x/base)
}

cces <-
  mutate(cces,
         age_bin = mfloor(age, 10),
         age_bin = ifelse(age < 20, 20, age_bin))
```


# Merge in reported turnout {.tabset}
I can't reconstruct a reported turnout from the cumulative file because the cumulative CCES has collapsed two categories of uncertain turnout: 1) reported vote but not validated by voter file and 2) turnout in the voter file not reported in the survey.


## 2008 turnout
```{r}
cces08 <-
  haven::read_dta("data/cces/cces_2008_common.dta") %>%
  dplyr::select(
    case_id = V100,
    turnout_report = CC403,
    turnout_valid  = vote_gen08) %>%

  mutate(case_id = as.character(case_id),
         year  = 2008,
         voted_report = as.numeric(turnout_report == 5 & !is.na(turnout_report)),
         voted_valid  = as.numeric(turnout_valid == 1)
         )
```

```{r}
setDT(cces08)

cces[cces08,
     `:=`(
       voted_report = i.voted_report,
       voted_valid  = i.voted_valid
     ),
     on = c("year", "case_id")
     ]
```


## 2012 turnout

```{r}
cces12 <-
  haven::read_dta("data/cces/CCES12_Common_VV.dta",
                  col_select = c("V101", "e2012g", "CC401", "CC410a",
                                 "CC422a", "CC422b")
  ) %>%
  mutate(
    year = 2012,
    case_id =  as.character(V101),
    voted_report = as.numeric(CC401 == 5 & !is.na(CC401)),
    # not "" or "MatchedNoVote"
    voted_valid  = as.numeric(e2012g %in%
                                c("Polling", "Early",
                                  "Absentee", "UnknownMethod", "Mail")),
    rr_overcame = 6 - as.numeric(CC422a),
    rr_slavery  = as.numeric(CC422b),
    racialres2  = cbind(rr_slavery, rr_overcame) %>%
            apply(2, scale) %>%
            { rowSums(.) / 2 }
  )

```

```{r}
setDT(cces12)

cces[cces12,
     `:=`(
       voted_report = i.voted_report,
       voted_valid  = i.voted_valid,
       racialres2      = i.racialres2
     ),
     on = c("year", "case_id")
     ]
```

## 2016 turnout
```{r}
cces16 <-
  haven::read_dta("data/cces/CCES16_Common_OUTPUT_Feb2018_VV.dta",
                  col_select = c("V101",
                                 "CC16_401", "CL_E2016GVM",
                                 "CC16_410a")
                  ) %>%

  mutate(
    year = 2016,
    case_id = as.character(V101),

    voted_report =  as.numeric(CC16_401 == 5 & !is.na(CC16_401)),
    # not "" or "unknown"
    voted_valid  =  as.numeric(CL_E2016GVM %in%
                                 c("polling", "unknown", "mail",
                                   "earlyVote", "absentee"))
  )

```


```{r}
setDT(cces16)

cces[cces16,
     `:=`(
       voted_report = i.voted_report,
       voted_valid  = i.voted_valid
     ),
     on = c("year", "case_id")
     ]
```

## 2020 turnout
```{r}
cces20 <-
  haven::read_dta("data/cces/CES20_Common_OUTPUT_vv.dta",
                  col_select = c(case_id = "caseid",
                                 voted_report = "CC20_401",
                                 vote_method  = "CL_2020gvm",
                                 vote_pres = "CC20_364b",
                                 "commonweight"
                                 )
                  ) %>%

  mutate(
    year = 2020,
    case_id = as.character(caseid),
    # 5 -> definitely voted
    voted_report = as.numeric(CC20_401 == 5 & !is.na(CC20_401)),
    # not "" or "unknown"
    voted_valid  = as.numeric(as_factor(CL_2020gvm, levels = "label") %in%
                                c("polling", "unknown", "mail",
                                   "earlyVote", "absentee"))
  )
```


```{r}
setDT(cces20)

cces[cces20,
     `:=`(
       voted_report = i.voted_report,
       voted_valid  = i.voted_valid
     ),
     on = c("year", "case_id")
     ]
```


# Code numeric vote choice
```{r}
cces <-
  mutate(cces,
         vote_pres_rep = as.numeric(vote_pres == "rep"),
         vote_pres_dem = as.numeric(vote_pres == "dem"),

         vote_pres3_report    = case_when(vote_pres == "dem" ~ -1L,
                                          vote_pres == "rep" ~  1L,
					  vote_pres == "other" ~ 0L,
                                          voted_report != 1 ~  NA_integer_),
         vote_pres3_valid    = case_when(vote_pres == "dem" ~ -1L,
                                         vote_pres == "rep" ~  1L,
					 vote_pres == "other" ~ 0L,
                                         voted_valid != 1 ~  NA_integer_))
```


# Code income quintile
```{r}
cces <-
  group_by(cces, year) %>%
  mutate(faminc_quin = ntile(fincome_ordinal, 5) %>%
           recode_factor(`1` = "1st",
                         `2` = "2nd",
                         `3` = "3rd",
                         `4` = "4th",
                         `5` = "5th"
           )
  ) %>% ungroup()
```

# Remake pid3 from pid7

Very strange discrepancy between ANES and CCES versions of the 3-level party ID variable. Creating our own out of the 7-leveled version, which is more consistent across surveys.

```{r}
cces <-
  mutate(cces,
         pid3 = recode_factor(pid7,
                              `strong dem`     = "dem",
                              `weak dem`       = "dem",
                              `indep dem`      = "dem",
                              `independent`    = "indep",
                              `other/not sure` = "other/not sure",
                              `indep rep`      = "rep",
                              `weak rep`       = "rep",
                              `strong rep`     = "rep",
                              )
         )

```



# Checks
```{r}
wmean <-
    function(x, na.rm = TRUE, w)
        wtd.mean(x = x, na.rm = na.rm, weights = w)

# Presidential years only
assert(cces, in_set(seq(2008, 2020, 4)), year) %>%
    # complete, categ. race
  assert(in_set(c("black", "white", "hispanic", "other")), race) %>%
  invisible()
```


```{r}
setDT(cces)

assert_that(
  cces[ , setequal(seq(2008, 2020, 4), unique(year))],

  # demographics -----
  # count should be white > black > hispanic
  cces[ , wmean(race == "white", w = weight) >
                wmean(race == "black", w = weight) &
              #
              # 2008 had 2000 Black and 2000 Hispanic respondents...
              wmean(race == "black", w = weight) >=
                wmean(race == "hispanic", w = weight),
            by = year]$V1 %>% all(),

  cces[ , between(wmean(gender == "female", w = weight), 0.45, 0.55)]
  )
```


No absurd election shares.
```{r}
group_assert(data = cces[voted_valid == 1L], by = "year",
             grp_expr = expr(   .(dem_vshare = wmean(vote_pres_dem, w = weight))  ),
             assertion = expr(   between(dem_vshare, 0.44, 0.53))   )

group_assert(cces[voted_valid == 1L], by = "year",
             grp_expr = expr(    .(rep_vshare = wmean(vote_pres_rep, w = weight))  ),
             assertion = expr(    between(rep_vshare, 0.38, 0.48))    )

# third party
group_assert(cces[voted_valid == 1L], by = "year",
             grp_expr = expr(   .(third_vshare = wmean(vote_pres_rep + vote_pres_dem == 0,
                                                       w = weight))),
             assertion = expr(    between(third_vshare, 0, 0.07))    )

# turnout
group_assert(cces, by = "year",
             grp_expr = expr(   .(turnout = wmean(voted_report == 1L, w = weight))   ),
             assertion = expr(    between(turnout, 0.60, 0.80))    )

group_assert(cces, by = "year",
             grp_expr = expr(   .(turnout = wmean(voted_valid == 1L, w = weight))   ),
             assertion = expr(    between(turnout, 0.4, 0.65))    )
```


# Export

Drop auxiliary columns.

```{r}
cces <- dplyr::select(cces, any_of(names(cces)),
                          faminc_quin, racialres2) %>%
            mutate_if(haven::is.labelled, as.character)

setDT(cces)
cces[ , year := as.character(year)]

saveRDS(cces, "data/cces_clean.rds")
```
